---
title: "Project Proposal"
author: "OutOfTheBoxplots"
output: github_document
editor_options: 
  chunk_output_type: console
---
```{r load-data, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
semester <- read_csv(here::here("data", "semester.csv"))
usage_chart <- read_csv(here::here("data", "usage_chart.csv"))
food_points <- read_csv(here::here("data", "input_food_points_data.csv"))
```

## High-Level Goal

The goal of this project is to build an R Shiny dashboard app that takes Duke 
students’ food plan data and generates a report on their usage.

## Project Goals

Every semester Duke offers the same various food plans to meet the diverse needs 
of their undergraduate students. These customized plans are tailored 
specifically for upper-class, first-year and off-campus students, yet most 
students, regardless of their food plan, face the same issue — they always run 
out of food points by the end of the semester. The dreaded feeling of bargaining 
for your friend’s spare food points is something all Duke students have felt at 
one point in time, and this is the motivation for a pertinent issue we would 
like to explore. With all the options Duke offers, why is it so hard to budget 
our food points, and what can help us budget better? In order to alleviate the 
mental hurdles that come with deciding how much money we can spend on food, we 
have decided to create a R Shiny dashboard app that uses a Duke student’s food 
plan and data from their DukeCard Account Transaction Report to generate a 
visualreport on their spending. To do this we will utilize Duke’s Fall 2021 and 
Spring 2022 Food Points Usage Chart. This chart breaks down all 8 of Duke’s food 
plans and the suggested usage rate based on daily and weekly averages. We plan 
to use this in combination with a student’s personal usage data, which includes 
the location food points were spent at — accessed through their [DukeCard eAccounts login](https://dukecard.duke.edu/manage-your-dukecard) — to provide a 
comprehensive report on their food spending. Students will generate a usage 
report from their online DukeCard statements. Students will then copy and paste this 
personal information into an Excel template we provide, to be uploaded to the R 
Shiny app. This Excel template can be found in our data folder and it looks like 
the following: 'input_food_points_data.csv'. 
Explicit instructions will be added to the beginning of the app that will
instruct the user about how to upload their data.

Using the user input template, we will create time series plot that depict user
food point data as lines over time by specific eatery or by all eateries. 
We will also include a map plot of campus with points placed over the image of 
Duke's campus, enlarged by the measure of frequency. 2 options for bar charts 
include a racing bar chart of a user's top 5 eateries over time or a bar chart
with a slider over time. We will also include the option for a user to chose 
the measure of comparison for eateries by frequency or spending. We will 
consider other visualizations as needed.


```{r glimpse-food-points, echo =FALSE}
glimpse(food_points)
```

We consulted Duke Dining on the best way to extract food point data and they 
believe our method is most efficient, rather than using an inaccessible api.

Students will then be provided information about how their spending compares to 
Duke's recommended spending by utilizing a trendline that shows their spending rate
for each week as the semester goes on, as well as other visualizations which 
communicate metrics about the times and eatery locations in which they spend the 
most food points. We hope that this tool can provide students with the necessary
insight on how much they can spend on food and where they should or shouldn’t 
spend their food, while staying within the limits of their food plan.

The first dataset we are using, `semester.csv`, offers a brief summary of each 
plan, which will allow us to quickly outline each plan’s offerings. Each 
observation is an offered Duke undergraduate food plan. The accompanying 
variables are information about the total number of points offered on the plan, 
the expected number of points to be spent per week, and the expected number of 
points to be spent per day. There are `r nrow(semester)` observations and 
`r ncol(semester)` variables in the data set. The data was collected from 
[Duke's Food Plan and Points Website](https://studentaffairs.duke.edu/dining/plans-points). 
The second dataset we are using, `usage_chart.csv`, offers in depth recommended 
spending information for each plan, which will allow us to offer recommendations 
to Duke students. Each observation is a week during a Duke undergraduate 
semester. The accompanying variables are information about the start date for 
the week in the Fall or Spring semester, and the expected number of food points 
to start that week with, depending on which plan a student has. There are 
`r nrow(usage_chart)` observations and `r ncol(usage_chart)` variables in the 
data set. The data was collected from the Fall 2021 and Spring 2022 Food Points 
Calculators found on [Duke's Food Plan and Points Website](https://studentaffairs.duke.edu/dining/plans-points). 
A comprehensive code book for each data set can be found in the README file 
inside the data folder in this project’s GitHub repository. A glimpse of each 
data set can be seen below:

`semester.csv`:
```{r glimpse-semester, echo = FALSE}
glimpse(semester)
```

`usage_chart.csv`:
```{r glimpse-usage-chart, echo = FALSE}
glimpse(usage_chart)
```

## Weekly Plan of Attack

*Week 3 (starting Nov 1): Complete Project Proposal*

- Finalize Project Repo, include code book and all necessary data to configure R 
Shiny app (Anna)
- Create plan of attack for the project and figure out roles (Blossom)
- Write up 1-2 paragraph description of our goals for project proposal (Hebron)
- Start testing R Shiny app dashboard and assist Anna with finalizing repo 
(Matthew)

*Week 4 (starting Nov 8): Update + Finalize Project Proposal*

- Create list of visualizations we want to see within app (Everyone)
  - Consider food points data over time, location, factor in student 
  breaks/holidays, potential finals season
- Add Sample Dataset (Matthew)
- Create Food Points Excel Template (Anna)
- Update project proposal based on peer feedback (Hebron)
- Start implementing visualization ideas (Everyone)

*Week 5-6 (starting Nov 15, starting Nov 22): Work on Data Visualizations*

- Create visualizations for data with focus over time (consider semester, 
student breaks, holidays, finals season) (Anna, Matthew)
- Create visualizations for data with focus by location (i.e. explore custom 
geospatial possibilities, bar charts) (Blossom, Hebron)

*Week 7 (starting Nov 29): Project Finishing Touches*

- Write Introduction (Everyone)
- Write Conclusion (Everyone)
- Contribute to Project write-up (Everyone)
- Practice Presentation (Everyone)
- Finalize Website and Presentation (Everyone)

## Organization of Project Repository

For our project repository organization we will have folders for: the proposal, 
the data, the dashboard, the presentation, and images. The dashboard folder will 
contain the script to create our final deliverable: the dashboard. The proposal 
folder will contain the rmd and md files for our proposal write up. Accompanying 
images that will be included in the dashboard will be placed in the images 
folder. The presentation folder will contain the code for a xaringan slideshow 
(if we choose to present using this format). The data folder will contain the 
csv files with information about Duke's food plans, as well as the 
Excel template that users of the dashboard can download along with example
datasets (we will decide on the best location for this template later on,
so it's exact location in the project repository is yet to be determined). 
Each folder contains a README file explaining in further detail the folder’s
contents, as well as a README file in the main folder, describing the overall
project. This overall README file will likely contain the finalized write-up,
too, and the accompanying rmd file will be placed in the main folder. 
